G\"{o}del's Ontological Proof of God's Existence (Draft)

### 2013-08-25 Natural Deduction/GodProof-ND.tex
Make the natural language statements in the ND proof agree with those in the abstract

\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}

@@

\section{Possible witnessing of positive properties}

\textbf{Axioms:}
\begin{itemize}
\item \textbf{(A1)} Properties necessarily entailed by \emph{positive} properties are also positive:
$$
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
$$
%
\item \textbf{(A2)} A property's negation is positive iff the property is not positive:
$$
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
$$
\end{itemize}

\noindent
\textbf{Lemma 1:} Positive properties possibly have a witness:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$

\noindent
\textbf{Formal proof:}

@@

\section{Possible existence of a God}

\textbf{Axioms:}
\begin{itemize}
\item \textbf{(A3)} Being God is a positive property:
$$
P(G)
$$
\end{itemize}

\noindent
\textbf{Lemma 2:} It is possible that a God exists:
$$
\pos \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

@@

\textbf{Definitions:}
\begin{itemize}
\item \textbf{(D1)} An individual is a \emph{God} if and only if he possesses all positive properties:
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\item \textbf{(D2)} A property is \emph{essential} for an individual if and only if it holds for that individual and necessarily entails every other property that holds for that individual: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))$
\end{itemize}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(A4)} Positive properties are necessarily positive:
$$
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
$$
\end{itemize}

\noindent
\textbf{Lemma 3}: If an individual is a God, then being God is an essential property for that individual:
$$
\all y.[G(y) \imp \ess{G}{y}]
$$

\noindent
\textbf{Formal proof:}

@@

Let the following derivation with the open assumption $G(x)$ be $\Pi_2$: 
\begin{prooftree}
       \AXC{$\psi(x)^1$} \RightLabel{}
             	\AXC{$\Pi_1 $}  \dashedLine
		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp$E}
       		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
		\AXC{$$} \RightLabel{Ax. 4}
       		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
		\UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp$E}  
          \BIC{$\Box P(\psi)$}\RightLabel{$\imp$I, 1} 
           \UIC{$\psi(x)\imp\Box P(\psi)$} 
\end{prooftree}

@@

\textbf{Definitions:}
\begin{itemize}
\item \textbf{(D3)} An individual \emph{necessarily exists} if and only if all its essential properties are necessarily witnessed:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)]
$$
\end{itemize}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(A5)} Necessary existence is a positive property:
$$
P(E)
$$
\end{itemize}

\noindent
\textbf{Auxiliary Lemma:} If there is a God, then there necessarily exists a God:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

@@

\section{Necessary existence of a God}

@@

\noindent
\textbf{Theorem:} The existence of a God is necessary:
$$
\nec \ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

@@

\section{God's existence}

\noindent
\textbf{Axioms:}
\begin{itemize}
\item \textbf{(M)} What is necessary is the case:
$$
\all \varphi. [\nec \varphi \imp \varphi]
$$
\end{itemize}

\noindent
\textbf{Corollary:} There exists a God:
$$
\ex x. G(x)
$$

\noindent
\textbf{Formal proof:}

>>>>

\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
\newcommand{\NE}[1]{\mathit{NE}(#1)}

@@

\section{Introduction}

\section{Natural Deduction}

\section{Possibly, God Exists}

\begin{description}
\item[A1] Either a property or its negation is positive, but not both:
$$
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
$$
%
\item[A2] A property necessarily implied by a positive property is positive:
$$
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)]
$$
\end{description}

\begin{description}
\item[T1] Positive properties are possibly exemplified:
$$
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
$$
\end{description}

@@

\begin{description}
\item[D1] A \emph{God-like} being possesses all positive properties:
$$
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
$$
\item[A3] The property of being God-like is positive:
$$
P(G)
$$
\end{description}

\begin{description}
\item[C1] Possibly, God exists:
$$
\pos \ex x. G(x)
$$
\end{description}

@@

\begin{description}
\item[A4] Positive properties are necessarily positive:
$$
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
$$
%
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))$
\end{description}

\begin{description}
\item[T2]: Being God-like is an essence of any
  God-like being:
$$
\all y.[G(y) \imp \ess{G}{y}]
$$
\end{description}

@@

Let the following derivation with the open assumption $G(x)$ be $\Pi_2$: 
\begin{prooftree}
       \AXC{$\psi(x)^1$} \RightLabel{}
             	\AXC{$\Pi_1 $}  \dashedLine
		\UIC{$\psi(x)\imp P(\psi)$}\RightLabel{$\imp$E}
       		\BIC{$P(\psi)$}\RightLabel{$\imp$E}  
		\AXC{$ $} \RightLabel{Ax. 4}
       		\UIC{$\forall \psi .(P(\psi)\imp \Box P(\psi))$}\RightLabel{$\forall$E}  
		\UIC{$P(\psi)\imp \Box P(\psi)$}\RightLabel{$\imp$E}  
          \BIC{$\Box P(\psi)$}\RightLabel{$\imp$I, 1} 
           \UIC{$\psi(x)\imp\Box P(\psi)$} 
\end{prooftree}

@@

\begin{description}
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
$$
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)]
$$
%
\item[A5] Necessary existence is a positive property:
$$
P(E)
$$
\end{description}

\begin{description}
\item[L1] If there is a God, then there necessarily exists a God:
$$
\ex z. G(z) \imp \nec \ex x. G(x)
$$
\end{description}

@@

\section{Necessarily, God exists}

@@

\begin{description}
\item[T3] The existence of a God is necessary:
$$
\nec \ex x. G(x)
$$
\end{description}

@@

\section{God exists}

\begin{description}
\item[M\phantom{0}] What is necessary is the case:
$$
\all \varphi. [\nec \varphi \imp \varphi]
$$
\end{description}

\begin{description}
\item[C2] There exists a God:
$$
\ex x. G(x)
$$
\end{description}
